Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-10 ms.tex
cleaned up3

\item Only for high X-ray radiative efficiencies similar to the maximum values predicted by ADAF models are the inferred accretion rates of quiescent early-type galaxies below the maximum value allowed for thermally-stable accretion $\dot{M}_{\rm TI}$ (Fig.~\ref{fig:miller}).  It is thus well possible that the current sample of X-ray detections is instead comprised mainly of nuclei undergoing non-steady accretion due to (local) thermal instabilities.  In this case there may exist a separate branch of low $L_X$ (undetected) thermally-stable nuclei, especially at low $M_{\bullet}$.  The existence of such a bimodal population would call into question extrapolating the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to constrain the SMBH occupation fraction in low mass galaxies (\citealt{Miller+15}).

@@

We conclude by drawing attention to the following limitations of our model.  First, by neglecting any large scale inflow of gas to the nucleus and assuming the only source of gas feeding the black hole is stellar wind mass loss from the local stellar population, our estimates represent a lower bound on the gas density of the CNM.  Our model therefore cannot account for most AGN activity.

The predicted SMBH accretion rate is also strongly tied to the age and radial distribution of the stellar population within the sphere of influence.  Our model cannot make definitive predictions for the accretion mode of individual galaxies without knowledge of their specific SFH.

We adopt halo average star formation rates measured in cosmological simulations by \citet{MosterNaab+:2013a}, and neglect any spatial variation in the star formation rate for any given galaxy. Galaxies of mass $\gsim 7\times 10^{10} \Msun$, assemble their stars from the inside out \citep{PerezCidFernandes+:2013a}, such that for a given halo mass the galactic nucleus may be older than in our model and thus more prone to thermal instability.

On the other hand, \citealt{LauerFaber+:2005a} find that $\sim
      60\%$ of cusp galaxies and $\sim29\%$ of core galaxies have
      (generally unresolved) emission in excess of the inward
      extrapolation of the Nuker law.  They note that these nuclei are bluer
      than the background galaxy starlight, suggesting the presence of
      younger stars towards galactic centre {\bf AG mass
        threshold for this effect}.

      It is thus possible that the mass injection to the nucleus could in some cases be dominated by a nuclear star cluster with its own break radius on parsec scales, i.e. much smaller than our fiducial value of $r_b \sim 100$ pc.  For a sufficiently high heating rate, a smaller break radius would not appreciably change the stagnation radius, resulting in a similar accretion rate onto the black hole.  On the other hand, a break radius smaller than the nominal stagnation radius would likely decrease the stagation radius and SMBH accretion rate.  
%      be for large break radii, the stagnation radius would move
 %     inwards somewhat. However note the for high heating rate ($v_w
  %    \gsim 500 $ km/s and small black hole masses $\Mbh \lsim 10^7
   %   \Msun$), we predict stagnation radii smaller than a parsec.

>>>>

\item  The accretion rates of early-type galaxies are below the maximum value for thermally stable flow, $\dot{M}_{\rm TI}$ (Fig.~\ref{fig:miller}), only for high X-ray radiative efficiencies, similar to the maximum values predicted by ADAF models.  Therefore the current X-ray detections could instead be comprised mostly of nuclei undergoing non-steady accretion due to thermal instabilities.  In this case there may exist a separate branch of low-$L_X$ (undetected) thermally-stable nuclei, especially at low $M_{\bullet}$.  Such a bimodal population would call into question methods of extrapolating the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ in order to constrain the SMBH occupation fraction in low mass galaxies (\citealt{Miller+15}).

@@

We conclude by drawing attention to the following limitations of our model.  First, by neglecting any large scale inflow of gas to the nucleus and assuming the only source of gas feeding the black hole is stellar wind mass loss from the local stellar population, our estimates represent a lower bound on the gas density of the CNM.  Our model therefore cannot account for AGN.

Predictions for the  SMBH accretion rate are strongly tied to the age and radial distribution of the stellar population within the sphere of influence.  Our model therefore cannot make definitive predictions for the accretion mode of individual galaxies without knowledge of their specific SFH.
We adopt halo average star formation rates measured in cosmological simulations by \citet{MosterNaab+:2013a}, and neglect any spatial variation in the star formation rate for any given galaxy. However, galaxies of mass $\gsim 7\times 10^{10} \Msun$ assemble their stars from the inside out \citep{PerezCidFernandes+:2013a}, resulting in galactic nuclei which may be systematically older than assumed in our model and thus more prone to thermal instability.

On the other hand, \citealt{LauerFaber+:2005a} find that $\sim
      60\%$ of cusp galaxies and $\sim29\%$ of core galaxies have
      (generally unresolved) emission in excess of the inward
      extrapolation of the Nuker law.  They note that these nuclei are bluer
      than the background galaxy starlight, suggesting the presence of
      younger stars towards galactic centre {\bf AG mass
        threshold for this effect}.  Mass injection to the nucleus could in some cases therefore be dominated by a nuclear star cluster with its own break radius on parsec scales, i.e. much smaller than our fiducial value of $r_b \sim 100$ pc.  In the case of high heating, a smaller break radius does not appreciably change the stagnation radius, resulting in a similar SMBH accretion rate.  For break radii smaller than the nominal stagnation radius, the presence of a smaller break radius would likely decrease the stagation radius and accretion rate.  
%      be for large break radii, the stagnation radius would move
 %     inwards somewhat. However note the for high heating rate ($v_w
  %    \gsim 500 $ km/s and small black hole masses $\Mbh \lsim 10^7
   %   \Msun$), we predict stagnation radii smaller than a parsec.
